Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation

### 2017-10-04 inst/doc/draft.Rmd
hypervolume update

---
title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India: vertical stratification reflects guild-dependent dynamics'
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: pdf_document
csl: mee.csl
bibliography: refs.bib
---

@@

functional trait; rainforest; dry season; forest strata; Western Ghats; ...

@@

Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975]. Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978]. Many theories have been proposed to explain how species coexistence can be maintained in ecological communities . Among them, two have been hotly debated and tested those last years. First niche theory states that the niche differences of species determine wether they can establish and coexist in communities [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that species coexistence is possible even when species are functionaly equivalent, under the sole influence of stochasticity in life, death, reproduction and dispersal [@hubbell_unified_2001]. Both theory can provide realistic expectations of taxonomic diversity [@chave_comparing_2002], and some recent approaches have advocated investigating patterns of functional [@westoby_land-plant_2006; @mcgill_rebuilding_2006] and phylogenetic diversity [@mouquet_ecophylogenetics:_2012] to trace the signature of ecological processes related to specific biological attributes and adaptations. 
These approaches have been applied to resolve the conundrum of high tree diversity in tropical forests [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012], but studies on these ecosystems still remain scarce. We here investigate taxonomic, functional and phylogenetic diversity in a low elevation rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], where few studies have assessed the relative influence of neutral and niche-based processes on local forest composition.

The variation in functional traits among species reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007]. These variation thus reflect fundamental trade-offs in ecological strategies along environmental gradients [@mark_westoby_plant_2002]. For instance, specific leaf area varies along a soil fertility-shade index and represent a trade-off between resource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to shared ecological strategies, functional traits will covary along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra. The leaf economics spectrum opposes productive ecological strategy to conservative persistent strategy. The wood economics spectrum distinguish species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, wiht high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009]. @baraloto_decoupled_2010 highlighted decoupled wood and leaf strategies allowing independant variation of these strategies, but @reich_world-wide_2014 conversely advocated that the two spectra have to integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing species to more conservative ones. Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest.

A major challenge in addressing community assembly is to relate local coexistence to a broader spatial context of species dynamics [@ricklefs_disintegration_2008; @munoz_neutral_2016]. Niche-based dynamics can entail community variation along broad environmental gradients, while local community dynamics can be neutral [@leibold_coexistence_2006]. Local community dynamics can further be influenced by niche differences under the influence of competition [@macarthur_limiting_1967]. In this perspective, @messier_trait_2016 have underlined that the significance of the leaf economics spectrum is basically scale-dependent, which can explain conflicting results across studies performed at different scales. In the same logic, a spatial hierarchy of ecological dynamics can entail a hierarchy of variation in functional traits from broad variation among species along large-scale environmental gradients, to intraspecific response to micro-environmental conditions at very fine scale [@albert_when_2011]. Therefore a multiscale framework is needed to represent the hierarchy of ecological processes from regional to local scale [@lortie_rethinking_2004]. Determining the extent of neutral and niche-based dynamics thus implies decomposing diversity patterns within and among communities [@ackerly_trait-based_2007], and designing appropriate statistical tests to disentangle the influence of local and larger-scale dynamics [@bernard-verdier_community_2012]. In this paper, we will decompose diversity patterns in a large rainforest plot into sources of variation within and between communities [@kraft_disentangling_2011], and investigate the fine-scale and plot-scale dynamics using appropriate null models. For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they challenged the existence of the leaf economics spectrum at community scale.

Apart from scale-dependent dynamics in space within and among communities, ecological guilds can coexist and occupy distinct niches within communities, while the species within guilds are ecologically similar [@wilson_guilds_1999; @vergnon_niches_2009]. The neutral theory has been basically applied in rainforests to adress diversity patterns in the guild of canopy trees [@hubbell_unified_2001], thus ignoring dynamics of understorey trees. However basic ecological constraints drive resource use strategies and biotic interactions in these ecosystems and imply considering functional strategies and relationship among guilds as well. Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988]. It thus provide an ideal context to analyze guild-dependent dynamics and ecological strategies in tropical rainforests. We will thus address diversity patterns and community assembly at guild-level.

In this paper, we used a trait-based approach to adress within-guild community dynamics within and among local tree assemblages in a tropical rain forest from Western Ghats, India. Although considered as a wet evergreen forest, the forest undergoes a long dry season and an intense rainy season under the influence of monsoon. In this context, we expect water stress and competition for resource to be the key drivers of forest dynamics. We thus adressed two main questions:

1. Do variations and covariations in functional traits reflect distinct ecological strategies within and among species ?
    i. How functional traits vary among species ?
    ii. Do covariations in functional traits reflect axes of ecological strategy ?
2. Does functional composition depends on potential local abiotic and biotic drivers of community dynamics ?
    i. Do local functional composition ($\alpha$ scale) reflect the influence of abiotic and biotic drivers of community dynamics ?
    ii. Do functional composition variation between communities ($\beta$ scale) reflect the changing of abiotic and biotic conditions ?

@@

The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996]. The plot is located near Uppangala village, and thus refers to as "Uppangala plot", in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India). Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996]. Climate and pedology of the 10-ha plot can be considered as homogeneous. The topography of the site is highly heterogeneous with a regular east-west alternation between thalwegs and ridges separated by steep slopes, so that topography can be assumed as a major source of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998]. We hypothesized that topographic variation should influence community variation depending on tree ecological strategies [@condit_spatial_2000].

## Floristic and trait data

Since 1989, all trees $\geq 30$ cm in girth at breast height (gbh) have been monitored in an area progressively extended to a contiguous 10-ha plot up to 2014. <!-- Need to add Cedric Vega in Acknowledgments for that. --> All trees were identified, tagged and mapped over the plot [@pelissier_twenty_2011]. 111 different species were recorded. We collected trait data on 89 over 111 (80.1%) species, which represented 99.0% of the tree individuals present in the Uppangala plot. Note that the 32 most abundant species, represent  more than 80% of Uppangala tree individuals. We sampled five individuals per species for these 32 species, and sampled one to five individuals from the remaining species (in total 267 individuals were sampled). For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected in consistent light conditions, development stage and sanitary state. Dawkins crown exposure index [@dawkins_management_1958] was recorder for each tree sampled. Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag untill their measurement in the following 24 hours. For some individuals, we were unable to collect an appropriate branch sample (11 missing species over 89 representing 1% of individuals present in the 10 ha plot).

@@

Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weights, area and thickness were measured, using a weighing scale, a bend scanner and a micrometer respectively. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$. Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias. The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues. Branch wood samples were sink in water to evaluate their volume by water displacement. Samples were then dried for 48 hours at 80 $^\circ C$ and weighted.

We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1). We choose those five traits because they have been widely used, in addition of there easier measurement, and they have been shown to be good indicator of species abilities to grow, survive and reproduce [@wilson_specific_1999].

Table 1: Functional traits measured. *Traits with their abbreviation, standard unit, and associated species strategy for functionning, survival and reproduction from @baraloto_decoupled_2010.*

```{r 1 Traits}
rm(list = ls()[-which(ls() == "cores")]); invisible(gc())
PFT <- genPFT()
table <- data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense vs. investment',
        'Leaf resource capture vs.investment',
        'Leaf defense vs. investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
)
kable(table, row.names = F)
rm(table)
```

@@

Because topography was assumed as a major source of environmental variation [@gimaret-carpentier_sampling_1998], we selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed using ArcGis [@desktop_arcgis_2011].

In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the environmental variables (see supplementary materials S1). We selected the less correlated variables having the most ecological meaning. Thus we reduced our explanatory variables to 4 variables: curvature representing the water flow paths and soil accumulation areas, wetness highlighting water accumulation areas, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on mutrient/water availability **ref**.

We selected two biotic factors related to local forest structure: canopy height and basal area. Canopy height, i.e. the maximum local height of the canopy, was calculated from a digital canopy model built from a LiDAR (Light Detection And Ranging) campaign done in 2013 over the Uppangala plot with a resolution of $1*1 m^2$. Basal area was calculated from 2013 and 2014 girth inventories of Uppangala plot on stems with a diameter above 30 centimeters at breast height ($DBH>30~cm$).

@@

To quantify trait variation whithin and between individuals and between species we calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values between leaves of the same individual (intra-individual), between trees of the same species (intra-specific for species with more than one individual), and between species (inter-specific). In addition, to assess the covariation of trait within species, and thus potential axes of strategy, we performed a principal component analysis (PCA) of the mean value of traits per species. Subsequent community-level analysis were based on trait values averaged per species.

### Community trait variability

Across the whole 10 hectare plot, $30*30 m^2$ quadrats were defined and represented 110 local communities. @pascal_wet_1988 defined four categories of strata groups: emergent below 39m, canopy tree between 25 and 39m, sub-canopy tree between 15 and 25m, and understorey tree below 15m. Due to topography, sub-canopy tree from Uppangala are often under light exposure and participates to canopy, especially in slopes. We defined two guilds to represent vertical stratification called canopy (C), composed of emergent, canopy and sub-canopy tree, and understorey (U) trees. The 110 local communities where split by those two guilds.

To characterize the local functional composition of canopy and understorey communities we used the community weighted mean (CWM), the community weighted variance (CWV), and the coefficient of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010]. CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $i$ species in the quadrat $j$, with each species, $i$, weighted by its relative abundance, $p_{i}$:
$$CWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
We also calculated the average of trait values based on species presence-absence, to further emphasize the contribution of less abundant and rare species in community trait composition [@cingolani_filtering_2007]. CWV for each trait $t$ was calculated as the mean distance of trait values, $t_{i}$, to $CWM$, of the $S$ species in the quadrat $j$, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - CWM_{j})^2 ~ (2)$$
CVNND for each trait $t$, was calculated as the standard deviation of the nearest neighbor distances (NND), $\sigma _{NND}$, of the $i$ individuals in each community, divided by the mean NND, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = {NND_{i} = min(d_{i,j})_{i \neq j}} ~ (3)$$

To investigate functional composition of communities we used a null model approach. A first null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix). The second model (NM2) randomized abundances among species within each community but kept the observed list of species and trait values inside communities [@bernard-verdier_community_2012]. Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic. For both CWV and CVNND, we performed a two-tailed test of the proportion of of communities differing from null expectation with adjusted p-values by Benjamini-Hochberg correction. CWV lower than expected under a null model represented convergence of trait values within the community, whereas CWV higher than null evidenced divergence in trait distribution [@bernard-verdier_community_2012]. CVNND was used to detect niche differentiation among coexisting species, which occured when the observed CVNND was lower than under a null model [@jung_intraspecific_2010]. We then calculated the standardized effect size (SES) as: $$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.
The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$). Analyses were conducted on the five functional traits and the species scores obtained on the 2 first PCA axes.

### Taxonomic and functional variation among communities

To investigate taxonomic and functional variation among communities we calculated $I_{st}$, for taxonomic composition and $U_{st}$ and $\tau _{st}$, for functional composition in abundance and presence-absence respectively [@hardy_characterizing_2007]. $I_{st}$ measures species turnover accounting for local abundance. $U_{st}$ integrate only functional turnover, weighting species by their local abundances; whereas $\tau _{st}$ express functional turnover, without accounting for species local abundances.

Variation in composition between community pairs with those metrics was translated as a taxonomic, functional and phylogenetic distance, and compared to the corresponding environmental and spatial distances. Environmental distances were calculated as euclidean distances between communities on the plan composed by the two first axes of environmental variables PCA (see supplementary material S1). Correlations with environment and space were assessed with two Mantel tests successively with taxonomic and functional, community composition. The firt Mantel test consisted in assessing the correlation between each metric distance matrix and the environmental distance matrix (E) and the second test, a partial Mantel test, assessed the correlation between each metric distance matrix and the spatial distance matrix knowing environmental distance matrix (E|S). This second test thus test the remaining spatial dependency in community composition once the influence of environmental variation is removed.

We used null models to test if the observed correlations differed significantly from expectations under null hypothesis. $I_{st}$ correlations with environment and space were tested with 999 repetions of the nul model described by @myers_beta-diversity_2013 which randomizes individuals among quadrat while preserving the regional abundance of species, and the number of individuals per community. $U st$ and $\tau st$ were tested by the nul model NM1 which randomizes traits among species.

All analyses were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
`ade4` [@dray_ade4_2007], 
`ape` [@paradis_ape:_2004], 
`phylosignal` [@keck_phylosignal:_2015], 
`spacodiR` [@eastman_spacodir:_2013],
and `vegan` [@oksanen_vegan:_2016].

@@

@@

We found a general increase of functional trait variation from intra-individual to inter-specific level (see table 2). Whereas LA shows already high variabililty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$). Finally the relatively low intra-individual variation ($cv_{intra}$ range from 0.06 to 0.09) allow us to consider the mean trait values per species for subsequent connumity-level analysis.

Table 2: Coefficient of variation of intra-individual, intra-specific and inter-specific trait values. *Traits with their range of values, mean and standard deviation, mean of coefficient of variation for leaves in all individuals (intra-individual), for trees in all species (intra-specific), and coefficient of variation for trees for all species. See table 1 for abbreviation.*

@@

## Functional composition of local communities

>>>>

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation
  rainforest of India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: word_document
csl: mee.csl
bibliography: refs.bib
---

@@

functional trait; rainforest; vertical forest strata; Western Ghats; environmental filternig; niche differentiation; ecological guild

@@

Understanding the determinants of species occurrence and co-occurrence is a long standing issue in ecology [@diamond_ecology_1975]. Tropical rainforests have especially fascinated ecologists due to their outstanding diversity [@connell_diversity_1978]. Niche theory explains species coexistence depending on niche differences to limit competitive exclusion [@weiher_assembly_1995; @lortie_rethinking_2004], while neutral theory shows that maintenance of high diversity is possible even for functionally equivalent species, under the influence of stochastic life, death, reproduction and dispersal dynamics [@hubbell_unified_2001]. Both theories can provide realistic expectations of taxonomic diversity [@chave_comparing_2002] so that inferring ecological processes from taxonomical composition patternis is challenging in community ecology.

Investigating patterns of functional or phylogenetic diversity may reveal signatures of ecological processes that are specific to the niche theory [@mcgill_rebuilding_2006; @westoby_land-plant_2006; @mouquet_ecophylogenetics:_2012]. Indeed, if taxonomic composition is expected to be spatially structured in both theories, due to the spatial structure of environment or to dispersal limitation, functional or phylogenetic assemblage are expected to be spatially structured only under the niche theory [@kraft_functional_2010; @baraloto_using_2012; @hardy_phylogenetic_2012; @parmentier_prevalence_2014]. We here investigate taxonomic and functional diversity in a topographically complex rainforest of the Western Ghats biodiversity hotspot [@myers_biodiversity_2000], India, where few studies have assessed the ecological drivers of local composition [@munoz2007estimating; @hardy_phylogenetic_2012,; @krishnadas2016environmental].

The variation in functional traits among species is expected to reflect different abilities to grow, survive and reproduce in a given environment [@violle_let_2007] and expresses fundamental trade-offs along environmental gradients [@mark_westoby_plant_2002]. For instance, the specific leaf area varies along a soil fertility-shade gradient and represents a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@hodgson_is_2011]. Fundamental trade-offs entail functional traits covariation along basic leaf [@wright_worldwide_2004 but see @osnas_global_2013 and @lloyd_photosynthetically_2013] and wood [@chave_towards_2009] economics spectra. The leaf economics spectrum opposes acquisitive to conservative ecological strategies, basically expressed at leaf level. **WES trade-off to clarify without CSR** The wood economics spectrum distinguishes species strategies over three axes spanning the basic CSR strategies of @grime_evidence_1977 : (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength [@chave_towards_2009]. @baraloto_decoupled_2010 found decoupled wood and leaf economics of trees in neotropical forest, but @reich_world-wide_2014 conversely posited that both spectra have to be integrated in a comprehensive whole plant economics spectrum [@freschet_evidence_2010] opposing fast growing acquisitive species to slow growing and more conservative ones. Here we address how varying wood and leaf ecological strategies determine species community assembly in an Indian rainforest.

A major challenge is to relate local coexistence in an ecological community to a broader context of species dynamics [@ricklefs_disintegration_2008; @munoz_neutral_2016]. The trait values of coexisting species represent a non-random sample of trait values found in a broader area, du to specific niche-based dynamics in the local community [@lortie_rethinking_2004; @bernard-verdier_community_2012]. Then the signature of local niche-based dynamics can be gasped by assessing the deviation of local community structure from a larger "species pool", classicaly using null models [@kraft_functional_2008]. Specifically, lower variance of functional traits than null is expected when a narrow range of strategies is locally successful [functional convergence, @ackerly_trait-based_2007], while larger variance and greater eveness is expected if biotic interactions entail niche differentiation [@macarthur_limiting_1967]. In addition, the nature of functional traits and ecological strategies driving species coexistence can change across scales. For instance, @messier_trait_2016 have underlined that the leaf economics spectrum basically expresses changes in ecological strategies along broad-scale environmental gradients, while it is not necessarily a primary driver of local coexistence. In this paper we will decompose the variation of tree functional traits in a large rainforest plot into sources of variation within communities [@kraft_disentangling_2011], using appropriate null models.

The concept of ecological community basically represents the spatial limits which species share resources and interact [**cite Loojien and van Andel 1999**], while the concept of guilds refer to groups of functionally similar species [@wilson_guilds_1999]. The two concepts are not exclusive, as ecological guilds can coexist and occupy distinct niches whithin communities [@vergnon_niches_2009]. In tropical forests, most emphasis has been put on the dynamics of the guilds of canopy trees [@hubbell_unified_2001] **Maxime disagree, Hubbel used 1 cm gbh**, but the way disctinct guilds of canopy and understorey trees coexist in communities is little addressed. Distinctive ecological constraints are expected to drive different resource use strategies and biotic interactions in these guilds, which should translate into different functional composition and response to environmental factors. Low elevation rainforests in South India are remarkably heterogeneous in vertical structure and encompass distinctive guilds across the vertical strata [@pascal_wet_1988]. It thus provides an appropriate context to adress guild-dependent dynamics and ecological strategies in tropical rainforests. We thus investigate here the diversity patterns and community assembly of rainforest trees at guild-level.

In this paper, we used a trait-based approach to address guild dynamics whithn local tree assembly along topographical gradients in a tropical rainforest from the Western Ghats, India. Although considered as wet evergreen, the forest undergoes contrasted long dry and intense rainy seasons under the influence of monsoon. In this context, we expected water stress and competition for resource to be primary drivers of forest dynamics and functional composition. A large forest plot (10ha) was subdivided in local elementary communities and we investigated how guild composition whithin communities varied with local environmental conditions. We addressed the following questions:

1. About the variation and covariation in functional traits within and among species,
    i. Does functional traits variation allow identifying distinctive ecological strategies among species ?
    ii. Does their covariations reflect basic dimensions of ecological strategy in the rainforest ?
2. About the role of functional trait variation in guild and community assembly,
    iii. Does the local functional composition $\alpha$ reflect the influence of guild and community dynamics ?

@@

The study was carried out in a 10-ha permanent plot established in 1990 by the French Institute of Pondicherry in an undisturbed, old-growth wet evergreen Dipterocarp forest [@pascal_structure_1996]. The plot is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), near the Uppangala village, and thus refers to as "Uppangala plot". Annual precipitation averages 5108 mm per year but is highly variable in time with heavy rainfall from June to October (90% of the rainfall during the monsoon) and a severe dry season (monthly rainfall $\leq 45$ mm) lasting 3.8 months [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27$^\circ C$, with a mean minimum about 25$^\circ C$ during July rainfall pick and a maximum mean temperature of 29$^\circ C$ in April [@pascal_structure_1996]. Climate and geology of the 10-ha plot can be considered as homogeneous. The topography of the site is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes, so that it is expected to be a major component of environmental variation within the forest plot [@gimaret-carpentier_sampling_1998]. Therefore, we hypothesized that topographic variation should influence community assembly and yield changing ecological strategies and functional composition [@condit_spatial_2000].

## Taxonomic composition and trait sampling

From 1989 to 2014, the trees $\geq 30$ cm in girth at breast height (gbh) were identified, tagged, mapped and monitored in an area progressively extended to a contiguous $300 * 330 m^2$ (ca. 10-ha) plot [@pelissier_twenty_2011]. The plot included 111 different tree species. We collected trait data from 89 over 111 (80.1%) species, which represented 99.0% of tree individuals present in the plot. Note that the 32 most abundant species alone, represented more than 80% of tree individuals. We sampled five individuals per species for these 32 species, and sampled one to five individuals for the remaining species (in total 267 individuals). For each individual we collected 5 mature shadow leaves and one branch sample with a diameter close to 1 cm following recommendation of @perez-harguindeguy_new_2013. Individuals and leaves were selected in consistent light conditions, development stage and sanitary state. Dawkins crown exposure index [@dawkins_management_1958] was recorded for each tree sampled. Leaf and branch samples were kept in humidified ziplock filled with enriched $CO_{2}$ air inside a black bag until measurement within the next 24 hours. For some individuals, we were unable to collect suitable branch sample (11 missing species over 89 representing 1% of individuals present).

@@

Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weights, thickness and area were measured, using a weighing scale (**add detail on brand and precision**), a micrometer (**add detail on brand and precision**) and a bend scanner with ImageJ software [**ref**] respectivelly. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 $^\circ C$. Once dried, the leaves were weighted directly after been extracted from the oven to avoid humidity bias. The bark and pit of branch samples were removed in order to keep only branch wood with no soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80 $^\circ C$ and weighted.

We calculated five functional traits of leaf and wood samples: leaf thickness (LT in $\mu m$ ), leaf area measured from scans (LA in $mm^2$), leaf dry matter content as the leaf dry weight divided by the leaf fresh weight (LDMC in $mg.g^-1$), specific leaf area as the leaf area divided by the leaf dry mass (SLA in $m^2.kg^-1$), and branch wood density  as the the wood sample dry weight divided by its volume (WD in $g.cm^-3$, see table 1). We choose those five traits because they proved to be good proxies of the major dimensions of ecological strategies in plants [@mark_westoby_plant_2002; @diaz_global_2015] and were reatively easy to measure on the field.

Table 1: Functional traits measured. *Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].*

```{r 1 Traits}
rm(list = ls()[-which(ls() == "cores")]); invisible(gc())
PFT <- genPFT()
table <- data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense and investment',
        'Leaf resource capture vs.investment',
        'Leaf defense and investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
)
kable(table, row.names = F)
rm(table)
```

@@

Topography was expected to be a major environmental factor influencing the assemby of tree communities [@gimaret-carpentier_sampling_1998]. We selected 8 topographical variables representing the abiotic environment: elevation, slope, curvature, plan curvature, profile curvature, wetness (function of both slope and elevation representing water flow and accumulation areas), southwesterness (SW, aspect in direction of south-west: $cos(aspect~in~degres + 225)$) derived from elevation data available within the plot using the RSAGA package [@brenning_statistical_2008], and distance to the nearest potential river computed in ArcGis [**ref**].

In order to identify the main sources of environmental variation and to avoid multi-colinearity in subsequent analyses, we performed a principal component analysis (PCA) of the abiotic environmental variables (see Supplementary Materials S1). We selected four less correlated variables: curvature representing local water flow paths and soil accumulation areas, wetness highlighting water accumulation areas over the whole study area, southwesterness representing aspect [@bunyan_effect_2015] and slope, which may have both a strong effect on forest dynamics and on nutrient/water availability **ref**.

In addition, we selected two biotic factors related to local forest structure: canopy height and basal area. Canopy height, i.e. the maximum local height of the canopy, was calculated from a 1-m resolution digital canopy model built from a LiDAR (Light Detection And Ranging) dataset acquired in 2013 over the Uppangala plot. Basal area was calculated the girth measurements taken within the 10-ha plot (census done in 2013 and 2014) for stems $>30~cm$ gbh.

@@

We calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among leaves of the same individual (intra-individual), among trees of the same species (intra-specific for species with more than one individual), and among species (inter-specific). We performed a principal component analysis (PCA) of the mean value of traits per species to identify main dimensions of ecological strategies along which traits can covary. Subsequent community-level analyses were based on trait values averaged per species.

### Guild-level analyses

$30*30 m^2$ quadrats were delimited in the 10 heactare plot and represented 110 local communities. @pascal_wet_1988 defined four vertical strata attained by adult trees: emergent above 39m, canopy trees between 25 and 39m, sub-canopy trees between 15 and 25m, and understorey trees below 15m. Despite the vertical hierarchy, steep topography could allow sub-canopy trees receiving much light, expecially in slopes, and thus participates to canopy. We defined two basic guilds called canopy (C), composed of emergent, canopy and sub-canopy species, and understorey (U) species. We analyze trait composition of these guilds in the 110 communities.

<!-- Metrics -->
We calculated weighted mean (GWM, where G stands for "guild"), weighted variance (GWV), and coefficient of variation of nearest neighbour distance (CVNND) [GWM and GWV, @violle_let_2007; CVNND, @jung_intraspecific_2010] for each guild in each community. In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of $i$ species weighted by species relative abundance $p_{i}$:
$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
where $S$ is the number of species in the quadrat.

We also calculated the average trait values based on species presence-absence, to addressthe influence less abundant and rare species on community trait composition [@cingolani_filtering_2007]. GWV of trait $t$ was calculated as :
$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~ (2)$$

CVNND of trait $t$ was calculated as the standard deviation of nearest neighbor distances (NND) $\sigma _{NND}$ divided by the mean NND $\mu _{NND}$; and NND was the minimal distance $d$ between a species and the most similar other species of the guild :
$$CVNND = \frac{\sigma _{NND}}{\mu _{NND}} ~ , ~ ~ ~ NND = [NND_{i} = min(d_{i,j})_{i \neq j}] ~ (3)$$

<!-- Null models -->
To assess significant changes in functional composition among guilds, we devised a null model apporoach. A first null model (NM1) randomized trait values among species in the whole plot (permuting raw in `Trait*Species` matrix). The second model (NM2) randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012]. **Spatial correlations in tree abundance are conserved limiting associated type I error.** GWM and CVNND metrics were compared to null model NM1 whereas GWV was studied with null model NM2 **To develop**.

<!-- Analysis -->
We performed 999 randomizations for each model and we derived a two-tailed test of the proportion of guilds deviating from the null expectation, with adjusted p-values by Benjamini-Hochberg correction **Citation**. Shifts in GWM values with changing environmental conditions could reflect the shorting of species strategies across assemblages [@ackerly_trait-based_2007]. GWV lower/higher than null represented convergence/divergence of trait values within the guild [@bernard-verdier_community_2012]. CVNND was used to detect niche differentiation among coexisting species, which occurred when observed CVNND was lower than null [@jung_intraspecific_2010].

To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution.

The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$). Analyses were conducted on the five functional traits and the species scores obtained on the 2 first PCA axes.

All analyses were conducted using the R 3.3.0  software [@r_core_team_r:_2016] with packages 
`ade4` [@dray_ade4_2007],
and `vegan` [@oksanen_vegan:_2016].

@@

**General statistics on guilds: size, richness...**

@@

We found overall increase of functional trait variation from intra-individual to inter-specific level (see table 2), even though LA showed almost as large variability at intra-individual than at intra-specific level ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$, wilcoxon rank test $p-value = 0.137$). Relatively low intra-individual variation ($cv_{intra}$ range from 0.06 to 0.09) allowed us considering mean trait values per species in subsequent analyses of the trait composition of guilds.

Table 2: Intra-individual, intra-specific and inter-specific component of trait variation. *The overall range, mean and standard deviation are provided for each trait. The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species. See table 1 for abbreviation.*

@@

## Functional composition of guilds
